---
title: "Tentative Topics Schedule"
author: |
  | MTH-201 Section E
  | Fall 2023
  | University of Portland
  | Alex John Quijano
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books & Online Resources Lists for the reading materials.
>
> The readings are not mandatory but it is encouraged.

<br>The "Reading" column in the table below contains page number on which it refers to a label in the Books & Online Resources List. For example "Pgs. 58-63 [CSV]" refers to pages 58-63 of the first item in the list, which is our main text book titled "Calculus: Single Variable (7th ed)".

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Activity** | **Homework** | **Lab** | **Reading** |
|---|---|---|---|---|---|---|
| **1** | M 8/28 | Introduction & Orientation | Function Placemats | - | Assigned: Lab 0 | Syllabus |
| - | W 8/30 | [1.7 Limits and Continuity](class-notes/1.7 Limits and Continuity.pdf){target="_blank"} | What's up with these functions?<br>[[Handout]](topics/1.7-limits-and-continuity/1.7_Limits_F23.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64ef88678586d2895a2aa247){target="_blank"} | Due: Precalculus Diagnostic Quiz | - | Pgs. 58-63 [CSV] |
| - | Th 8/31 | [1.8 Extending the Idea of a Limit](class-notes/1.8 Extending the Idea of a Limit.pdf){target="_blank"} | Limits and Continuity<br>[[Handout]](topics/1.8-limits-and-continuity/11.8_Limits_F23.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64ef88678586d2895a2aa247){target="_blank"} | - | - | Pgs. 67-72 [CSV] |
| - | F 9/1 | [1.9 Limits and Algebra](class-notes/1.9 Limits and Algebra.pdf){target="_blank"} | Computing Limits<br>[[Handout]](topics/1.9-limits-and-algebra/1_9Limits.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64ef88678586d2895a2aa247){target="_blank"} | Due: Section 1.7 | - | Pgs. 75-80 [CSV] |
| **2** | M 9/4 | [2.1 Measuring Speed](class-notes/2.1 Measuring Speed.pdf){target="_blank"} | How do we measure speed?<br>[[Handout]](topics/2.1-measuring-speed/2.1 Speed.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64f545408b96bd2220eedfc1){target="_blank"} | Due: Section 1.8 | - | Pgs. 84-88 [CSV] |
| - | W 9/6 | [2.2 Derivative at a Point](class-notes/2.2 Derivative at a Point.pdf){target="_blank"} | Derivative at a Point<br>[[Handout]](topics/2.2-derivative-at-a-point/2.2 Derivative.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64f8c7741f18f956129bd287){target="_blank"} | Due: Section 1.9 | - | Pgs. 91-95 [CSV] |
| - | Th 9/7 | Lab 1: Limits and Continuity | - | - | Assigned: Lab 1 | - |
| - | F 9/8 | **Proficiency Quiz<br>L1, L2\*** | - | Due: Section 2.1 | **Due: Lab 0** | - |
| **3** | M 9/11 | [2.3 Derivative Function Preview](class-notes/2.3 Derivative Function Preview.pdf){target="_blank"} | Derivative Function Preview<br>[[Handout]](topics/2.3-derivative-function-preview/2.3 Derivative Function Preview.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64fe787caaf2700e78da672b){target="_blank"} | Due: Section 2.2 | - | Pgs. 99-104 [CSV] |
| - | W 9/13 | [2.3 Derivative Function](class-notes/2.3 Derivative Function.pdf){target="_blank"} | The Derivative Function<br>[[Handout]](topics/2.3-derivative-function/2.3 Derivative Function.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64fe7f73f0257dc5eed274d4){target="_blank"} | - | - | - |
| - | Th 9/14 | [2.4 Interpreting the Derivative](class-notes/2.4 Interpreting the Derivative.pdf){target="_blank"} | Interpretations of the Derivative<br>[[Handout]](topics/2.4-interpreting-the-derivative/2.4 Interpreting the Derivative.pdf){target="_blank"}<br>[[Desmos]](https://student.desmos.com/activitybuilder/student-greeting/64fe7f73f0257dc5eed274d4){target="_blank"} | - |  | Pgs. 108-110 [CSV] |
| - | F 9/15 | **Proficiency Quiz<br>L2\* (2nd), L3** | - | Due: Section 2.3 | **Due Lab 1** | - |
| **4** | M 9/18 | [2.5 Second Derivative](class-notes/2.5 Second Derivative.pdf){target="_blank"} | Concavity<br>[[Handout]](topics/2.5-second-derivative/2.5 Concavity.pdf){target="_blank"} | Due: Section 2.4 | - | Pgs. 115-119 [CSV] |
| - | W 9/20 | [2.6 Differentiability](class-notes/2.6 Differentiability.pdf){target="_blank"} | Differentiability<br>[[Handout]](topics/2.6-differentiability/2.6 Differentiability.pdf){target="_blank"} | Due: Section 2.5 | - | Pgs. 123-126 [CSV] |
| - | Th 9/21 | Lab 2: Corona Virus Cases | - | - | Assigned: Lab 2 | - |
| - | F 9/22 | **Proficiency Quiz<br>D1\*, D2** | - | Due: Section 2.6 | - | - |
| **5** | M 9/25 | [3.1 Powers and Polynomials](class-notes/3.1 Powers and Polynomials.pdf){target="_blank"} | Derivatives of Polynomials<br>[[Handout]](topics/3.1-powers-and-polynomials/3.1 Derivatives of Polynomials.pdf){target="_blank"} | - | - | Pgs. 130-136 [CSV] |
| - | W 9/27 | [Review: Exponent Properties](class-notes/Review Exponent Properties.pdf){target="_blank"} | Exponent and Log Review<br>[[Handout]](topics/3.2.0-review-exponent-properties/3.2.0 Exponent Review.pdf){target="_blank"} | Due: Section 3.1 | - | - |
| - | Th 9/28 | [3.2 The Exponential Function](class-notes/3.2 The Exponential Function.pdf){target="_blank"} | Derivatives of Exponential Functions<br>[[Handout]](topics/3.2-the-exponential-function/3.2 Exponential Functions.pdf){target="_blank"} | - | - | Pgs. 140-142 [CSV] |
| - | F 9/29 | **Proficiency Quiz<br>D1\* (2nd), D3** | - | Due Exponent Review | **Due: Lab 2** | - |
| **6** | M 10/2 | [3.3 Product and Quotient Rule](class-notes/3.3 Product and Quotient Rule.pdf){target="_blank"} | Product and Quotient Rules<br>[[Handout]](topics/3.3-product-and-quotient-rule/3.3 Product and Quotient Rules.pdf){target="_blank"} | Due: Section 3.2 | - | Pgs. 144-147 [CSV] |
| - | W 10/4 | [3.4 Chain Rule](class-notes/3.4 Chain Rule.pdf){target="_blank"} | Chain Rule<br>[[Handout]](topics/3.4-chain-rule/3.4 Chain Rule.pdf){target="_blank"} | Due: Section 3.3 | - | Pgs. 151-155 [CSV] |
| - | Th 10/5 | [Review: Trigonometry](class-notes/Review Trigonometry.pdf){target="_blank"} | Trigonometry Review<br>[[Handout]](topics/3.5.0-review-trigonometry/Trig Review.pdf){target="_blank"} | - | - | - |
| - | F 10/6 | **Proficiency Quiz<br>D4, D5\*** | - | Due: Section 3.4 | - | - |
| **7** | M 10/9 | [3.5 Trig Functions](class-notes/3.5 Trig Functions.pdf){target="_blank"} | Derivatives of Trigonometric Functions<br>[[Handout]](topics/3.5-trig-functions/3.5 Trig Derivatives.pdf){target="_blank"} | Due: Trig Review<br>Due: Inverses Review | - | Pgs. 158-162 [CSV] |
| - | W 10/11 | [3.6 Inverse Functions](class-notes/3.6 Inverse Functions){target="_blank"} | Derivatives of Inverse Functions<br>[[Handout]](topics/3.6-inverse-functions/3.6 Inverse Functions.pdf){target="_blank"} | Due: Section 3.5 | - | Pgs. 164-167 [CSV] |
| - | Th 10/12 | *Canceled* | - | - | Assigned: Lab 3 | - |
| - | F 10/13 | **Proficiency Quiz<br>D5\* (2nd), D6** | - | Due: Section 3.6 | - | - |
| **8** | M 10/16 | _Fall Break_ | - | - | - | - |
| - | W 10/18 | _Fall Break_ | - | - | - | - |
| - | Th 10/19 | _Fall Break_ | - | - | - | - |
| - | F 10/20 | _Fall Break_ | - | - | - | - |
| **9** | M 10/23 | [3.7 Implicit Differentiation](class-notes/3.7 Implicit Differentiation.pdf){target="_blank"} | Implicit Differentiation<br>[[Handout]](topics/3.7-implicit-differentiation/3.7 Implicit Differentiation.pdf){target="_blank"} | Due: Derivative Practice | - | Pgs. 171-173 [CSV] |
| - | W 10/25 | *Mid-semester Reflection* | Informal Student Evaluations of Teaching Survey | Due: Section 3.7 | - | Pgs. 178-182 [CSV] |
| - | Th 10/26 | [3.9 Linear Approximation](class-notes/3.9 Linear Approximation.pdf){target="_blank"} | Linear Approximation<br>[[Handout]](topics/3.9-linear-approximation/3.9 Linear Approximation.pdf){target="_blank"} | - | - | Pgs. 192-198 [CSV] |
| - | F 10/27 | **Proficiency Quiz<br>D7, D8\*** | - | Due: Section 3.9<br>Due: Section 3.10 (Bonus) | **Due: Lab 3** | - |
| **10** | M 10/30 | [4.1 Using Derivative](class-notes/4.1 Using Derivative.pdf){target="_blank"} | Using First and Second Derivatives<br>[[Handout]](topics/4.1-using-derivative/4.1 Using First and Second Derivatives.pdf){target="_blank"} | Due: Section 4.1 | - | Pgs. 203-208 [CSV] |
| - | W 11/1 | [4.2 Optimization Part 1](class-notes/4.2 Optimization Part 1.pdf){target="_blank"} | Optimization Part 1<br>[[Handout]](topics/4.2-optimization-part-1/4.2 Optimization Part 1.pdf){target="_blank"} | - | - | Pgs. 212-218 [CSV] |
| - | Th 11/2 | Lab 4: Global Warming | - | - | Assigned: Lab 4 | - |
| - | F 11/3 | **Proficiency Quiz<br>D8\* (2nd), A1, A2\*** | - | Due: Section 4.2 | - | - |
| **11** | M 11/6 | [4.3 Optimization Part 2](class-notes/4.2 Optimization Part 2.pdf){target="_blank"} | Optimization Part 2<br>[[Handout]](topics/4.2-optimization-part-2/4.2 Optimization Part 2.pdf){target="_blank"} | - | - | - |
| - | W 11/8 | 4.3 Optimization Part 2 Cont. | - | Due: Section 4.3 | - | Pgs. 224-229 [CSV] |
| - | Th 11/9 | [4.6 Related Rates](class-notes/4.6 Related Rates.pdf){target="_blank"} | Related Rates<br>[[Handout]](topics/4.6-related-rates/4.6 Related Rates.pdf){target="_blank"} | - | - | Pgs. 243-246 [CSV] |
| - | F 11/10 | **Proficiency Quiz<br>A2\* (2nd), A3\*** | - | Due: Section 4.4 (Bonus) | **Due: Lab 4** | - |
| **12** | M 11/13 | [4.6 Related Rates Cont.](class-notes/4.6 Related Rates Cont.pdf){target="_blank"} | - | - | - | - |
| - | W 11/15 | [4.7 L'Hopital](class-notes/4.7 L'Hopital's Rule.pdf){target="_blank"} | L'Hopital's Rule<br>[[Handout]](topics/4.7-lhopital-rule/4.7 LHopital.pdf){target="_blank"} | Due: Section 4.6 | - | Pgs. 252-256 [CSV] |
| - | Th 11/16 | [5.1 Distance Traveled](class-notes/5.1 Distance Traveled.pdf){target="_blank"} | Measuring Distance Traveled<br>[[Handout]](topics/5.1-distance-traveled/5.1 Measuring distance traveled.pdf){target="_blank"} | - | - | Pgs. 272-278 [CSV] |
| - | F 11/17 | **Proficiency Quiz<br>L4, A3\* (2nd), A4** | - | Due: Section 4.7<br>Due: Section 4.8 (Bonus) | - | - |
| **13** | M 11/20 | 5.2 Definite Integral | Definite Integral<br>[[Handout]](topics/5.1-definite-integral/5.2 Definite Integral.pdf){target="_blank"} | Due: Section 5.1 | - | Pgs. 283-288 [CSV] |
| - | W 11/22 | Lab 5: Applications | - | Due: Section 5.2 | Assigned: Lab 5 | - |
| - | Th 11/23 | _Thanksgiving Vacation_ | - | - | - | - |
| - | F 11/24 | _Thanksgiving Vacation_ | - | - | - | - |
| **14** | M 11/27 | 5.3 Fundamental Theorem of Calculus Part 1 | The Fundamental Theorem and Interpretations<br>[[Handout 1]](topics/5.3-fundamental-theorem-of-calculus-part-1/5.3 Warm-up.pdf){target="_blank"}<br>[[Handout 2]](topics/5.3-fundamental-theorem-of-calculus-part-1/5.3 Net Change Theorem.pdf){target="_blank"} | - | - | Pgs. 292-297 [CSV] |
| - | W 11/29 | 5.4 Theorems About Integrals | Theorems About Definite Integrals<br>[[Handout]](topics/5.4-theorems-about-integrals/5.4 Theorems about Definite Integrals.pdf){target="_blank"} | Due: Section 5.3 | - | Pgs. 302-309 [CSV] |
| - | Th 11/30 | 6.1 Antiderivatives (Numerically & Graphically) | Antiderivatives Graphically<br>[[Handout]](topics/6.1-antiderivatives-numerically-and-graphically/6.1 Antiderivatives Graphically.pdf){target="_blank"} | - | - | Pgs. 316-317 [CSV] |
| - | F 12/1 | **Proficiency Quiz<br>I1, I2\*, I3\*** | - | Due: Section 5.4 | **Due: Lab 5** | - |
| **15** | M 12/4 | 6.2 Antiderivatives (Analytically) | TBA | Due: Section 6.1<br>Due: Section 6.2 | - | Pgs. 322-326 [CSV] |
| - | W 12/6 | 6.4 Fundamental Theorem of Calculus Part 2 | TBA | - | - | Pgs. 335-338 [CSV] |
| - | Th 12/7 | Review | - | - | - | - |
| - | F 12/8 | **Proficiency Quiz<br>I2\* (2nd), I3\* (2nd)** | - | Due: Section 6.3 (Bonus)<br>Due: Section 6.4 | - | - |
| **16** | Tu 12/12 | **Final Exam<br>Tu 10:30 AM - 12:30 PM** | - | - | - | - |

<br>

## **Books & Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

### Main Textbook

[CSV] `r capture.output(print(refs["hughshallet2017"]))`

<br>
